Tidy data structure to support exploration and modeling of temporal-context data

### 2018-05-30 main.Rmd
added one paragraph for index

---
title: "Tidy data structure to support exploration and modeling of temporal-context data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: 'Temporal-context data is often rich with information and time formats, for example multiple observational units, different time lengths, heterogeneous data types, nested and crossed factors and etc. This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and modelling routines. Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntatical and user-oriented approach in which it imposes nested or crossed structures on the data. This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context. A sound pipeline practice facilitates a succinct and transparent workflow for analyzing temporal data. Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage. The architecture of tidy temporal data has been implemented in the R package **tsibble**.'
keywords: "temporal context, time series, data structure, R"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: no
    number_sections: yes
    citation_package: biblatex
---

@@

Time representation: year, year-quarter, year-month, year-week, dates, date-times.

For data indexed in regular time space, the time interval is obtained by computing the greatest common divisor from positive time distances in a data table. This suggests that each observational unit collected at the same interval forms a table.

>>>>

---
title: "Tidy data structure to support exploration and modeling of temporal-context data"
author:
  - familyname: Wang
    othernames: Earo
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : earo.wang@monash.edu
    correspondingauthor: true
  - familyname: Cook
    othernames: Dianne
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : dicook@monash.edu
  - familyname: Hyndman
    othernames: Rob J
    address: Department of Econometrics and Business Statistics, \newline Monash University, VIC 3800 \newline Australia.
    email : rob.hyndman@monash.edu
abstract: 'Temporal-context data is often rich with information and time formats, for example multiple observational units, different time lengths, heterogeneous data types, nested and crossed factors and etc. This work presents a cohesive and conceptual framework for organizing and manipulating temporal data, which in turn flows into visualization and modelling routines. Tidy data principles are applied and extended to temporal data: (1) mapping the semantics of a dataset into its physical layout, (2) an explicitly declared index variable representing time, (3) a "key" comprised of single or multiple variables to uniquely identify units over time, using a syntactical and user-oriented approach in which it imposes nested or crossed structures on the data. This tidy data representation most naturally supports thinking of operations on the data as building blocks, forming part of a "data pipeline" in time-based context. A sound pipeline practice facilitates a succinct and transparent workflow for analyzing temporal data. Applications are included to illustrate tidy temporal data structure, data pipeline ideas and usage. The architecture of tidy temporal data has been implemented in the R package **tsibble**.'
keywords: "temporal context, time series, data structure, R"
wpnumber: no/yr
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
output:
  MonashEBSTemplates::workingpaper:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: no
    number_sections: yes
    citation_package: biblatex
---

@@

For data indexed in regular time space, the time interval is obtained by computing the greatest common divisor based on the non-negative differences of a given time index, although the data may include implicit time gaps. This implies that all the units in the table must share a common interval. It is relatively easy to determine if the data is daily or sub-daily since most statistical packages have built-in time representations for date-times and dates, for example their corresponding classes `POSIXct` and `Date` in R. However these turn out inappropriate for weekly, monthly and quarterly data. For example, the smallest time unit of monthly data is months in a year, which is equispaced points in a year. However, using the first day of the month can range from 28 to 31 days, which is irregularly spaced. A genuine year-month representation should avoid associating days with its format. Displaying the year and month components suffice to signal its monthly aggregated values instead of observed values at a particular timestamp.
